\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-05 papers/methods/paper.tex
Taking Dillon's comments on board

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. {\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}

In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae. {\green Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as we will required reduced systematic uncertainties to fully utilised increase statistics} \citep{Betoule2014, Scolnic2017}.

As such, staggering effort is being put into developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.  \citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of many factors to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.

@@

where $M_B$ is the mean absolute magnitude for all SN Ia  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by rearranging equation \eqref{eq:standard} to isolate $\mu(z)$. The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.

@@

where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Betoule2014} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
{\green 
\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right],\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
\end{align}
}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size. Whilst binning data results in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

@@

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and physical motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus easily incorporate different distributions, populations, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.

With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision. Their model was validated to be free of significant biases using fits to thirty realisations of supernova datasets. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}. For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.

@@

The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below. Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models and increasing simulation independence to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.

@@

We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections. We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}. The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.

@@

For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty. However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$. When using this simplification, our latent variables are given by
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
\end{align}
As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including intrinsic dispersion in the supernova model.

@@

\subsubsection{Type Ia population} 
\label{sec:underlying}
The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude. We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift. Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.

@@

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works. However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof of concepts we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates we leave for future improvements during the analysis of the DES photometric sample.

@@

With the {\green reported summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}. We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model. We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.

@@

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations. This extra dispersion is modelled with redshift independence.

@@

\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times \hat{m}. \label{eq:l2}
\end{align}

@@

This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --  the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes

@@

\begin{align}
d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
\end{align}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{efficiency.pdf}
	\end{center}
	\caption{The efficiency for supernova discovery at an arbitrary redshift. Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution. The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude. The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated. This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellowz}.}
	\label{fig:efficiency}
\end{figure}

@@

The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function). Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis. Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.

@@

Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys, where the LowZ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}. Simulation specifics can be found in \citet{Kessler18}.

In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based. We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.

>>>>

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. {\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}

In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae. {\green Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as we will required reduced systematic uncertainties to fully utilised increase statistics} \citep{Betoule2014, Scolnic2017}.

As such, {\green considerable resources are} being put into developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.  \citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of many factors to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.

@@

where $M_B$ is the mean absolute magnitude for all SN Ia  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by rearranging equation \eqref{eq:standard} to isolate $\mu(z)$. The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Scolnic2017}.

@@

where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Betoule2014} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
{\green 
\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}},
\end{align}
}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size. Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

@@

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and {\green physically} motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus easily incorporate different distributions, populations, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.

With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is determined perfectly. Their model was validated to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}. For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.

@@

The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below. Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models. {\green We increase simulation independence by not utilising bias-corrections} to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.

@@

We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, {\green treatment of an increased number of systematics}, and analytic correction of selection effects, including systematic uncertainty on those corrections. We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}. The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.

@@

For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty. However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$. When using this simplification, our latent variables are given by
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
\end{align}
As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including {\green full treatment of} intrinsic dispersion in the supernova model.

@@

\subsubsection{Type Ia population} 
\label{sec:underlying}
The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude. We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift. Each set of nodes are anchored to equally span the {\green redshift range }of each survey. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ {\green that} allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors. We sample $\delta_c$ in log space for efficiency in sampling close to zero.

@@

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works. However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernova falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof-of-concept we provide an accurate probability estimate. Further investigation on the effect of inaccurate {\green estimates will be left for} future improvements during the analysis of the DES photometric sample.

@@

With the {\green fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}. We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model. We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.

@@

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations. This magnitude dispersion is modelled with redshift independence.

@@

\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot \hat{m}. \label{eq:l2}
\end{align}

@@

This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --  the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. {\green A comprehensive explanation of all systematics is given in \citet{Brout18SYS}.} This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes

@@

\begin{align}
d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{align}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{efficiency.pdf}
	\end{center}
	\caption{The efficiency for supernova discovery at an arbitrary redshift. Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution. The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude. The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated. This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
	\label{fig:efficiency}
\end{figure}

@@

The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function). Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis. Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.

@@

Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}. Simulation specifics can be found in \citet{Kessler18}.

In addition to the improvements in the scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based. We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.
